MS ANXIETY

### 2025-01-23 scripts/MSAnxiety_manuscript_JAMA_revision_20250116.Rmd
Add files via upload

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(tidygam)
library(psych)
#library(tidymv)
library(aod)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/Library/CloudStorage/Box-Box/BBL/ms_generic_scripts/msdepression_functions.R")
```

@@

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller. It does however highlight the

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

@@

#### Anxiety dose
  - anxiety dose correlates with disease burden (total brain volume )
```{r lesion_volumes_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

@@

 #mean UF, depDiagnosisMS+noDepression   -0.05029    0.03028  -1.661   0.0978 . 
dep_mean_UF_gam_prop <- gam(mean_UF_prop ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_UF_gam_prop)

@@

###### Lesion & brainvol by depGroupVar (lesion vol - Y, tbv - N, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the lesion and brain vol findings are nonspecific t so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean vol mimosa lesions, depDiagnosisMS+noDepression 
#(T = -3.164, P = 0.002, Cohen's f2 =  0.128, Effect size = small)
dep_mean_lesion_vol_gam_prop <- gam(volume_of_mimosa_lesions ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_lesion_vol_gam_prop)

@@

  #mean brain vol mimosa lesions, depDiagnosisMS+noDepression  
 #(T = 0.016, P = 0.988, Cohen's f2 =  0.333, Effect size = medium)
dep_mean_total_brain_vol_gam_prop <- gam(total_volume ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_total_brain_vol_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_total_brain_vol_gam_prop)

@@

 ##### healthy vs sick - prop
#Fornix gam Diagnosis; DiagnosisMS+Anxiety    0.03690    0.05405   0.683   0.4961    - NS
anxietyDiagnosis_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)

@@

  #gam, Fornix anxiety_dose   0.02303    0.02401   0.959    0.338    - NS
  anxietyGrouping_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam_prop)

@@

###### JAMA ANALYSIS- UF Anxiety diagnosis + lesion vol
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

  #how about in fornix

  corr_LV_to_fornix <- cor.test(uncinate_by_dx_df$volume_of_mimosa_lesions,uncinate_by_dx_df$mean_fornix_volume) #corr 0.69; 
  print(corr_LV_to_fornix) #0.3452654

  lesion_vol_fornix_lm <- lm(uncinate_by_dx_df$volume_of_mimosa_lesions~ uncinate_by_dx_df$mean_fornix_volume)
  summary(lesion_vol_fornix_lm) #T = 4.047

  ## Try scaling mean volume
 # disease_mean_scaled_UF_vol_gam <- gam(mean_UF_vol_scaled_total_vol_lesions ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
  #summary(disease_mean_scaled_UF_vol_gam) #- N.S.

  #print summary stats; (T = -0.556, P = 0.579, Cohen's f2 =  0.007, Effect size = minimal)"
  #gam_summary_stats(disease_mean_scaled_UF_vol_gam)

## modeling this doesn't make sense
#disease_mean_UF_prop_w_LV_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F) + volume_of_mimosa_lesions, data = uncinate_by_dx_df)
#summary(disease_mean_UF_prop_w_LV_gam)

 #print summary stats
 # gam_summary_stats(disease_mean_UF_prop_w_LV_gam)
```
###### JAMA ANALYSIS-  UF Anxiety dose + lesion vol
```{r uncinateByAnxietyDose}

@@

```
###### JAMA ANALYSIS- UF Anxiety diagnosis + total brain vol
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

@@

###Promis data in UF - groups too small
#### Uncinate by anxiety diagnosis (N.S) 
```{r uncinate_by_dx}

### Results:  
#define group
  uncinate_by_dx_df <- df_people_with_promis_data_scores_added_unique %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=19

#bilateral UF prop ~ Diagnosis: DiagnosisMS+Anxiety -0.02490    0.11510  -0.216   0.8317 
disease_mean_UF_prop_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_prop_gam)

    #print summary stats
 gam_summary_stats(disease_mean_UF_prop_gam)

#### Uncinate by Anxiety Dose (N.S.)
```{r uncinateByAnxietyDose}

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_people_with_promis_data_scores_added_unique %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=64

  #bilat anxiety dose -anxiety_dose   -0.006261   0.055025  -0.114 0.909793       NS
  anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

    #print summary stats

 gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

```{r ants}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(aod)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/Library/CloudStorage/Box-Box/BBL/msanxiety/scripts/FINAL/ms_functions.R")
```

@@

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller.

@@

#### Anxiety dose
  - anxiety dose correlates with disease burden (total brain volume )
```{r tbv_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

@@

 #mean UF, depDiagnosisMS+noDepression  T = -1.659, P = 0.098, Cohen's f2 =  0.105
dep_mean_UF_gam_prop <- gam(mean_UF_prop ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F) + total_volume, data = df_dep)
summary(dep_mean_UF_gam_prop)

@@

###### Lesion & brainvol by depGroupVar (lesion vol - Y, tbv - N, N.S)
```{r UF_by_depGroupVar_lesion_tbv}
# Specificity - want to make sure the lesion and brain vol findings are nonspecific t so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean vol mimosa lesions, depDiagnosisMS+noDepression 
#(T = -3.2, P = 0.002, Cohen's f2 =  0.178, Effect size = medium)
dep_mean_lesion_vol_gam_prop <- gam(volume_of_mimosa_lesions ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F) + total_volume, data = df_dep)
summary(dep_mean_lesion_vol_gam_prop)

@@

  #mean brain vol mimosa lesions, depDiagnosisMS+noDepression  
 #(T = 0.016, P = 0.988, Cohen's f2 =  0.333, Effect size = medium)
#dep_mean_total_brain_vol_gam_prop <- gam(total_volume ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
#summary(dep_mean_total_brain_vol_gam_prop)

  #print summary stats
 #gam_summary_stats(dep_mean_total_brain_vol_gam_prop)

@@

 ##### healthy vs sick - prop
#Fornix gam Diagnosis; DiagnosisMS+Anxiety    (T = 0.683, P = 0.496, Cohen's f2 =  0.024, Effect size = small)
anxietyDiagnosis_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T) + total_volume, data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)

@@

  #gam, Fornix anxiety_dose   (T = 0.959, P = 0.338, Cohen's f2 =  0, Effect size = minimal)"
  anxietyGrouping_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T) + total_volume, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam_prop)

@@

###### JAMA ANALYSIS- UF Anxiety diagnosis + lesion vol
```{r uncinate_by_dx_lesion_vol}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

```
###### JAMA ANALYSIS-  UF Anxiety dose + lesion vol
```{r uncinateByAnxietyDose_lesion_vol}

@@

```
###### JAMA ANALYSIS- UF Anxiety diagnosis + total brain vol
```{r uncinate_by_dx_tbv}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

##### JAMA ANALYSIS- correction for multiple comparisons
```{r sensi_correction}
#uf dep group
#fornix by anxiety diag
#fornix by anxiety dose
#lesion vol anxiety diag
#lesion vol anxiety dose

anxiety_dx_vol_mimosa_lesions = summary(volume_of_mimosa_lesions_gam)$p.table[2,4]
anxiety_dose_vol_mimosa_lesions = summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)$p.table[2,4]
dep_mean_lesion_vol_gam = summary(dep_mean_lesion_vol_gam_prop)$p.table[2,4]
dep_mean_gam = summary(dep_mean_UF_gam_prop)$p.table[2,4]
anxiety_dx_fornix = summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)$p.table[2,4]
anxiety_dose_fornix = summary(anxietyGrouping_mean_F_by_sex_gam_prop)$p.table[2,4]

p_vals <- array(c(anxiety_dx_vol_mimosa_lesions,
                 anxiety_dose_vol_mimosa_lesions,
                 dep_mean_lesion_vol_gam,
                 dep_mean_gam,
                 anxiety_dx_fornix,
                 anxiety_dose_fornix))

#0.005324721 0.004567230 0.004567230 0.147282059 0.496139621 0.405577736
p_val_fdr <- p.adjust(method = "fdr", p_vals)
print(p_val_fdr)

@@
